---
title: "p8105_hw3_bes2161.Rmd"
author: "Brittany Shea"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyr)
library(lubridate)
library(ggridges)
library(patchwork)
library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Description of the dataset

The dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Each row represents an Instacart order. Variables include user, product, and order which includes day, time, and days since previous order. Product names, department, aisle, and order history are also included. There are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Here is a table that summarizes how many items were ordered from aisles. There are a total of 134 aisles, and the most ordered items are ordered from aisles with fresh vegetables and fresh fruits.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Here is a plot with the number of items ordered in each aisle, in ascending order by number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Here is a table which shows the top three popular items in each of the aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Here is a 2 x 7 table that shows the mean hour of the day when Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. Pink Lady Apples are purchased earlier in the day than Coffee Ice Cream, except for day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2

#### Read in and clean the data 

```{r}
accel_data = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count") %>%
  mutate(wday_wend = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday"))
```

This dataset includes 50,400 rows and 6 total columns. The variables include the week number, day id, day of the week, activity count for each minute of a 24-hour day starting at midnight, and whether the day is during the week or on a weekend. 

#### Table showing total activity variable for each day

```{r}
accel_data %>% 
  group_by(day_id) %>% 
  summarize(sum(activity_count)) %>% 
  knitr::kable(digits = 2)
```

It appears that activity decreases toward the beginning of the week; for example, Tuesdays appear to be a day with less activity compared to other days. However, activity seems to be higher on the weekends. 

#### Single-panel plot showing the 24-hour activity time courses for each day

```{r}
accel_data %>% 
  group_by(day_id) %>% 
  ggplot(aes(x = minute, y = activity_count)) + 
  geom_line(aes(color = day))
```

This graph shows there are peaks and dips in activity levels throughout the day. Activity increases in the beginning of the day, then decreases, and picks up again towards the end of the day. It also appears that activity is generally higher on weekends compared to weekdays. 

### Problem 3

### Read in the data

```{r}
data("ny_noaa")
```

### Description of the data

This dataset includes data on five variables for all NY weather stations from January 1, 1981-December 31, 2010. There are 2,595,176 rows and 7 total columns. The dataset includes variables such as id, date, precipitation, snowfall, snow depth, and maximum and minimum temperatures. There are several missing values (NAs) for each of these variables.

### Clean the data

```{r}
ny_noaa <- ny_noaa %>%
  janitor::clean_names() %>%
  mutate(tmax = as.numeric(tmax),
         tmin = as.numeric(tmin),
         prcp = as.numeric(prcp),
         snow = as.numeric(snow),
         snwd = as.numeric(snwd)) %>%
  mutate( 
    tmin = tmin / 10,
    tmax = tmax / 10,
    prcp = prcp / 10) %>%
  separate(date, c("Year", "Month", "Day"))
```

#### Most commonly observed values for snowfall

```{r}
ny_noaa %>%
  group_by(snow) %>%
  summarize(n()) %>%
  top_n(2)
```

For snowfall, the most commonly observed values are 0 and NA as weather stations do not collect all of the same variables. Some of the stations only report precipitation data. Additionally, snow does not fall for several months of the year.

### Two-panel plot showing the average max temperature in January and July in each station across years

```{r}
ny_noaa %>%
  group_by(Month, Year, id) %>%
  filter(Month %in% c("01", "07")) %>%
  summarize(tmax_mean = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = tmax_mean, group = id)) +
  geom_line(alpha = 0.5, color = "blue") + facet_grid(~ Month)
```

The average max temperature in January is lower than that in July in each station across years. There are some outliers, for example the temperature in January dropped lower during some years than other years, and some years the temperature dropped comparatively lower in July. It also appears that the average max temperature in January is increasing slightly.

#### Two-panel plot showing tmax vs tmin for the full dataset 

```{r}
ny_noaa %>%
  group_by(tmax, tmin) %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_line(alpha = 0.5, color = "red")
```

#### Plot showing distribution of snowfall values greater than 0 and less than 100 separately by year

```{r}
ny_noaa %>% 
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = snow, fill = Year)) +
  geom_density(alpha = .5) +
  theme(legend.position = "none") %>%
  labs(
    title = "Yearly Snowfall",
    x = "Snowfall",
    y = "Year"
  )
```
